---
title: "hw2"
author: "Yu Si"
date: "10/9/2021"
output: github_document
---
```{r}
library(tidyverse)
library(readxl)
library(haven)
```

##Problem 1
```{r}
# Read and clean the Mr. Trash Wheel sheet
trash_data = read_excel("./Trash-Wheel-Collection-Totals-7-2020-2.xlsx", col_names = TRUE, range = "A2:N534")%>% 
  janitor::clean_names() %>% 
  drop_na(dumpster) %>% 
  mutate(sports_balls = round(sports_balls))

#Read and clean precipitation data for 2018 and 2019. 
data_2018 = read_excel("./Trash-Wheel-Collection-Totals-7-2020-2.xlsx", sheet = 7, range = "A2:B14") %>% 
  drop_na() %>% 
  mutate(Year = "2018")
data_2019 = read_excel("./Trash-Wheel-Collection-Totals-7-2020-2.xlsx", sheet = 6, range = "A2:B14") %>% 
  drop_na() %>% 
  mutate(Year = "2019")

#Combine precipitation datasets and convert month to a character variable 
data_2018_2019 = bind_rows(data_2018, data_2019) %>% 
  mutate(Month = month.name[Month])

total_2018 = sum(pull(data_2018, Total))
trash_2019 = filter(trash_data, year == "2019")
sports_balls_median_2019 = median(pull(trash_2019, sports_balls))
```

In the trash_data, there are 344 rows and 14 columns. There are "dumpster","month","year", "date", "weight_tons", "volume_cubic_yards", "plastic_bottles", "polystyrene", "cigarette_butts", "glass_bottles", "grocery_bags", "chip_bags", "sports_balls", "homes_powered" observations. 
In data_2018, there are 6 rows and 3 columns. There are "Month" "Total" "Year" observations. 
In data_2019, there are 12 rows and 14 columns. There are "Month" "Total" "Year" observations. 
the total precipitation in 2018 is 70.33.
the median number of sports balls in a dumpster in 2019 is 9.
